Title

### 2013-04-09 TCAD2013/overclocking_tcad.tex
Modified CSA, revised by David, v1

\section{Carry Select Adder}
\subsection{Introduction}
Since the delay of RCA is determined by the length of carry chain, there are several alternative adder architectures proposed to shorten the carry chain. For instance, the carry select adder (CSA) is designed to overlap carry propagation in sections such that the operating speed can be boosted. In a CSA, the carry chain is divided into multiple stages. Each stage contains two RCAs and two multiplexers. For a given input, computations are performed twice simultaneously, with the carry input as zero and one respectively. The results are then selected according to the actual carry input. Although this structure brings timing benefits, it costs extra hardware resources than RCA as the carry chain is duplicated and multiplexers are expensive especially in the FPGA technology. In this section, silicon area is incorporated as the third metric besides accuracy and performance. We explore the trade-offs between these three factors for both adder structures. The objective is to demonstrate the best design choice for a given set of design constraints.

@@

\subsection{Timing Models for Carry Select Adder}
We initially describe the modelling method for the CSA timing, with the aim of forming the relationship between the operating frequency and the corresponding maximum word-length of CSA. This information can then be employed to determine the truncation error based on the models presented in Section.xxx.

In a CSA with $s$ stages~($s\geqslant 2$), let the stage delay be denoted by $d_{0}\dots d_{s-1}$, where $d_{0}$ and $d_{s-1}$ represent the delay of the most significant stage and the least significant stage, respectively. Unlike other stages, the least significant stage is built by one RCA only, since it is directly driven by the carry input. In our analysis, we follow the previous assumption that the critical path delay is due to carry propagation as well as, in this case, multiplexing the carry output. Therefore, the delay of the $i^{th}$ stage can be obtained through~(\ref{CSA_SingleStageDelay}), where $\mu_{c}$ denotes the delay of $1$-bit carry propagation and $\mu_{mux}$ denotes the delay of multiplexing.
%Note that the delay of the ${(s-1)}^{th}$ and ${(s-2)}^{th}$ stages are identical, because the least significant stage of CSA is built by RCA.
\begin{eqnarray}\label{CSA_SingleStageDelay}
  d_i=\left\{
    \begin{matrix}
      n_i\cdot \mu_{c}+(i+1)\cdot\mu_{mux}, &\textrm{if $i\in\left[0,s-2\right]$}\\
      d_{s-2}, & \textrm{if $i=s-1$}
    \end{matrix}
  \right.
\end{eqnarray}

Under the timing-driven design environment, the delay of each stage of CSA is equalized in order to achieve the fastest operation. That is,~(\ref{CSA_DelayEqual}) can be obtained.
\begin{eqnarray}\label{CSA_DelayEqual}
  d_0=d_1=\cdots=d_{s-1}
\end{eqnarray}

This design metric also consumes the maximum word-length of CSA for a given frequency, potentially corresponds to the minimum precision loss at inputs. Substituting (\ref{CSA_SingleStageDelay}) into (\ref{CSA_DelayEqual}) yields (\ref{CSA_DelayRep}), where $n_i$ denotes the CSA word-length of stage $i$.
%Note that the word-length of stage $s-1$ and $s-2$ are identical, since the least significant stage is composed of RCA.
\begin{eqnarray}\label{CSA_DelayRep}
 n_i=\left\{
	\begin{matrix}
	  n_0-i\cdot\frac{\mu_{mux}}{\mu_c}, & \textrm{if $i\in\left[0,s-2\right]$}\\
	  n_{s-2}, &\textrm{if $i=s-1$}
      %n_0-(c-2)\cdot\frac{\mu_{mux}}{\mu_c}, & \textrm{if $i=c-1$ and $c\geqslant2$}
	\end{matrix}
    \right.
\end{eqnarray}

The word-length of CSA is then given by (\ref{CSA_WL}).
\begin{eqnarray}\label{CSA_WL}
  \begin{split}
    n_{CSA} &=\sum_{i=0}^{s-1}n_{i} \\
    &= s\cdot n_{0}-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+1)(s-2)}{2}
  \end{split}
\end{eqnarray}

Conventionally for a given frequency constraint, the word-length of both RCA and CSA should be properly selected to avoid timing violation, as presented in~(\ref{CSA_RCA}), where $n_{RCA}$ is determined by current timing information through (xxx).
%In the conventional situation, the word-length of both RCA and CSA should be properly selected in order to meet timing. In this case,~(\ref{CSA_RCA}) can be obtained, where $n_{RCA}$ is determined by current timing information through (xxx).
\begin{eqnarray}\label{CSA_RCA}
%  \mu_{c}\cdot n_{RCA}=\mu_{c}n_0+\mu_{mux}
  \mu_{c}\cdot n_{RCA}=d_i
\end{eqnarray}

Combining~(\ref{CSA_WL}) and~(\ref{CSA_RCA}), we derive the representation of the word-length of CSA in terms of a given timing constraint, as presented in~(\ref{CSA_Timing}).
% we form the relationship between the word-length of CSA and RCA under a given timing constraint, as presented in~(\ref{CSA_Timing}).
\begin{eqnarray}\label{CSA_Timing}
  n_{CSA}=s\cdot n_{RCA}-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+2)(s-1)}{2}
\end{eqnarray}

It can be seen that $s=1$ leads to $n_{CSA}=n_{RCA}$. This is because RCA forms the least significant stage of CSA. In addition, combine (\ref{CSA_DelayRep}) and (\ref{CSA_RCA}) to ensure $n_{s-1}>0$, we obtain the upper bond of the stage number by~(\ref{CSA_StageNumber}).
\begin{eqnarray}\label{CSA_StageNumber}
  s<n_{RCA}\cdot\frac{\mu_c}{\mu_{mux}}+1
\end{eqnarray}
%the word-length of each CSA stage is allocated to ensure that

\subsection{Model Verification}
As seen in (\ref{CSA_Timing}), the value of $\mu_{mux}/\mu_c$ should be determined before applying the timing model. This ratio can be obtained through timing analysis instead of experiments. Specifically, varying the word-length of a single stage $n_i$ and recording the corresponding delay value $d_i$ from the timing analysis tool. According to~(\ref{CSA_SingleStageDelay}), the ratio can be obtained by fitting those values using linear polynomials.

Using this information, we verify our timing model with experimental results, which are obtained from post place and route simulations on Xilinx Virtex-6 FPGAs. Fig.~\ref{CSA Model Verification} illustrates both the modeled value and the experimental results of the maximum word-length of the 2-stage CSA for various operating frequencies. In the following experiments within this section, the input data are randomly sampled from a 16-bit data set, which follows uniform distribution. It can be seen that the the modeled value match well with the experimental results.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=3in]{./Figures/Model.eps}
  \caption{Comparison of the maximum word-length for given operating frequencies between the modeled value and the experimental results from FPGA simulations.}
  \label{CSA Model Verification}
\end{figure}

@@

\subsection{Area Overhead}
We then compare different adder structures in terms of the maximum word-length with respect to different operating frequencies and the corresponding area consumption. Fig.~\ref{CSA3stage Timing} illustrates the maximum word-lengths for RCA and CSA with 2 stages and 3 stages under a range of operating frequencies. We only investigate 3 stages as the maximum stages number predicted in (\ref{CSA_StageNumber}) is 3. However, similar analysis can be performed on CSA with any stage numbers and the overall conclusion will still hold. It can be seen from Fig.~\ref{CSA3stage Timing} that in comparison to RCA, CSA achieves greater word-length when frequency is initially increased. This corresponds to smaller truncation errors. RCA only outperforms than CSA when very high frequency is applied. In addition, the word-length of 3-stage CSA is always greater than 2-stage CSA across the entire frequency domain, as expected.
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/CSA3stage_Timing.eps}
%  \caption{Maximum word-length for RCA and CSA across a variety of frequencies.}
%  \label{CSA3stage Timing}
%\end{figure}

\begin{figure}[tbp]
    \begin{minipage}[b]{.48\textwidth}
        \centering
        \subfigure[Maximum word-length.]{
            \includegraphics[width=3.3in]{./Figures/CSA3stage_Timing.eps}
            \label{CSA3stage Timing}
        }
        \subfigure[Hardware resource usage.]{
            \includegraphics[width=3.3in]{./Figures/CSA3stage_Area.eps}
            \label{CSA3stage Area}
        }
    \end{minipage}
    \centering
    \caption{Comparison between RCA and CSA across a variety of frequency values in terms of the maximum word-length of input signal and the area consumption.}
\end{figure}

However, the accuracy benefits brought by CSA comes with the cost of large area overhead. Fig.~\ref{CSA3stage Area} depicts the number of Look-Up Tables (LUTs) in the FPGA used for all three structures. It can be seen that in order to meet a given frequency, the 3-stage CSA consumes $2.4\times\sim3.7\times$ area than RCA, while the number of the 2-stage CSA is $1.7\times\sim3.1\times$. This finding poses a question of whether CSA with higher stage numbers still offer better accuracy than RCA under a limited area budget.
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/CSA3stage_Area.eps}
%  \caption{Area overhead for RCA and CSA.}
%  \label{CSA3stage Area}
%\end{figure}

\subsection{Exploring Trade-offs Between Accuracy, Performance and Area}
To answer this question, the silicon area is incorporated into the trade-offs together with accuracy and performance. If the available hardware resources are limited, the full word-length of both CSA and RCA might not be implemented. The precision loss might results in large errors even at low frequencies. This is demonstrated by experiments where area constraints are applied besides the timing requirement. In the conventional design scenario, both of the constraints are met by reducing the word-length of the input signal. The errors at the outputs are recorded with the reference to the original word-length, which is 16 bits in the following experiments. In addition, we propose another circuit designing method where RCA is implemented with the maximum possible word-length under the given area budget, while the timing constraints are met by overclocking.

Our results could be of interest to a circuit designer in two aspects. On one hand, it is desirable to create a circuit that can run at a given frequency with the minimum achievable output errors and resource usage. On the other hand, the algorithm designer will wish the circuit to operate as fast as possible with the minimum area whilst a certain error budget can be tolerated. We select the optimum design methodology for both situations.

In the first case, for a certain pair of $\{Area, Frequency\}$ constraint, the optimum design method is selected based on the following criteria:
%For instance, Fig.xxx to Fig.xxx illustrate error expectations of all the described design scenarios when setting the LUTs number to 45, 35, 25 and 15 respectively. The optimal design metric with the minimum error expectation is labelled for each area constraint. For a given frequency and area constraint, the optimum design is chosen based on the following metrics:
\begin{itemize}
  \item Design with the best accuracy at outputs is the optimum design;
  \item If multiple designs achieve the same accuracy, then the design with minimum area is the optimum design;
  \item If the accuracy and area are identical for multiple designs, they are all treated as the optimum design.
\end{itemize}

The accuracy here is assessed in terms of the mean error at outputs. The optimum design method with respect to a variety of operating frequencies and area consumptions is demonstrated in Fig.~\ref{Tradeoff}, from which several observations can be made. If the available area is large enough, CSA serves as the optimum design option because it inherently operates faster than RCA, and it can be implemented with full precision with unlimited hardware resources. The 2-stage CSA is better than the 3-stage CSA when frequency is initially increased, as it consumes less area although both of them achieve the same error expectation. For a tighter area budget, only part of the CSA can be implemented, whilst RCA still keeps full precision. In this case, area instead of timing becomes the dominate factor. This leads to precision loss for CSA from the start. Therefore we see RCA with overclocking is the optimum design method across almost the whole frequency domain. For an even stringent area constraint, the word-length of RCA is also limited. This results in truncation error initially for all design scenarios. Meanwhile, CSA with high stage numbers might not even be implemented at all.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=3.5in]{./Figures/Tradeoff.eps}
  \caption{Demonstration of the optimum design methodology which achieves the minimum error at outputs with respect to a variety of frequency and area constraints.}
  \label{Tradeoff}
\end{figure}

Similarly in the second case where a pair of $\{Area, Accuracy\}$ constraint is applied, a certain amount of error can be tolerated when designing circuits. The error specifications are evaluated in terms of mean relative error (MRE), as presented in~(\ref{MRE}), where $E_{error}$ and $E_{out}$ refer to the mean value of error and the mean value of outputs, respectively.
\begin{eqnarray}\label{MRE}
  MRE=\left|\frac{E_{error}}{E_{out}}\right|\times 100\%
\end{eqnarray}

In our experiments, MRE is set ranging from $0.001\%$ to $50\%$. For a certain MRE, design with the maximum operating frequencies is labeled as the optimum design. Moreover, the smallest design is the optimum one if multiple structures operate at the same frequency, with a certain area. Based on these criteria, the decision graph is depicted in Fig.~(\ref{Tradeoff_Error}).
%We also notice that for small frequencies, RCA with either overclocking or truncation of word-length of the input signal is the best design choice for any area constraint, since
\begin{figure}[htbp]
    \centering
    \includegraphics[width=3.5in]{./Figures/Tradeoff_Error.eps}
    \caption{Demonstration of the optimum design methodology which runs at the fastest frequency with respect to a variety of accuracy and area constraints.}
    \label{Tradeoff_Error}
\end{figure}

For a tight accuracy requirement, i.e. $MRE<0.005\%$, CSA serves as the optimum design choice with respect to large accessible area, as it intrinsically operates faster than RCA. When the area budget shrinks, RCA exceeds because the precision is CSA is limited. Similarly the precision of RCA is reduced when critical area requirement is applied.

However, if the tolerable error increases, i.e. $MRE>1\%$, the overclocked RCA achieves the fastest operating frequencies with respect to most area constraints.

>>>>

\section{Carry Select Adder}
\subsection{Introduction}
Since the delay of RCA is determined by the length of carry chain, the carry select adder (CSA) is designed to shorten the carry chain by overlapping carry propagation in sections such that the operating speed can be boosted. In a CSA, the carry chain is divided into multiple stages, as seen in Fig.~xxx. Each stage contains two RCAs and two multiplexers. For a given input, two computations are performed simultaneously where the carry input is zero and one respectively. One of these two results is then selected according to the actual carry input. Although this structure brings timing benefits, it costs extra hardware resources compared to a standard RCA because the carry chain is duplicated. Furthermore, in FPGA technology, multiplexers are expensive. Due to this reason, we explore the trade-offs between silicon area, accuracy and performance of these two adder structures in this section.

%, silicon area is incorporated as the third metric besides accuracy and performance. We explore the trade-offs between these three factors for both adder structures. The objective is to demonstrate the best design choice for a given set of design constraints.

@@

\subsection{Timing Models for Carry Select Adder}
We initially model the CSA in order to understand the relationship between the operating frequency and the maximum word-length of the CSA. This information can then be employed to determine the truncation error based on the models presented in Section.xxx.

In a CSA with $s$ stages~($s\geqslant 2$), let the stage delay be denoted by $d_{s-1},~\dots,~d_{0}$, where $d_{s-1}$ and $d_{0}$ represent the delay of the most significant and the least significant stages, respectively. In our analysis, we follow the assumption that the critical path delay is due to carry propagation and multiplexing the carry output. Note that unlike other stages, the least significant stage is only built by one RCA without multiplexers, since it is directly driven by the carry input. Hence we obtain the delay of the $i^{th}$ stage as presented in~(\ref{CSA_SingleStageDelay}), where $\mu_{c}$, $\mu_{mux}$ and $n_i$ denote the delay of $1$-bit carry propagation, the delay of multiplexing and the word-length of the $i^{th}$ stage of the CSA, respectively.
\begin{eqnarray}\label{CSA_SingleStageDelay}
  d_i=\left\{
    \begin{matrix}
      n_i\cdot \mu_{c}+(s-i)\cdot\mu_{mux}, &\textrm{if $i\in\left[1,s-1\right]$}\\
      n_0\cdot \mu_c + (s-1)\cdot\mu_{mux}, & \textrm{if $i=0$}
    \end{matrix}
  \right.
\end{eqnarray}

Under the timing-driven design environment, the delay of each stage of CSA is equalized in order to achieve the fastest operation, as presented in~(\ref{CSA_DelayEqual}).
\begin{eqnarray}\label{CSA_DelayEqual}
  d_{s-1}=d_{s-2}=\cdots=d_{0}
\end{eqnarray}

In this case, combining~(\ref{CSA_SingleStageDelay}) and (\ref{CSA_DelayEqual}) yields $n_i$, which is represented by the word-length of the most significant stage $n_{s-1}$, in (\ref{CSA_DelayRep}).
%This design metric also consumes the maximum word-length of CSA for a given frequency, potentially corresponds to the minimum precision loss at inputs. Substituting (\ref{CSA_SingleStageDelay}) into (\ref{CSA_DelayEqual}) yields (\ref{CSA_DelayRep}), where $n_i$ denotes the CSA word-length of stage $i$.
\begin{eqnarray}\label{CSA_DelayRep}
 n_i=\left\{
	\begin{matrix}
	  n_{s-1}-(s-1-i)\cdot\frac{\mu_{mux}}{\mu_c}, & \textrm{if $i\in\left[1,s-1\right]$}\\
	  n_{s-1}-(s-2)\cdot\frac{\mu_{mux}}{\mu_c}, &\textrm{if $i=0$}
	\end{matrix}
    \right.
\end{eqnarray}

Summing up $n_i$ gives the total word-length of the CSA in (\ref{CSA_WL}).
%The total word-length of the CSA is the sum of every stage, as given by (\ref{CSA_WL}).
\begin{eqnarray}\label{CSA_WL}
  \begin{split}
    n_{CSA} &=\sum_{i=0}^{s-1}n_{i} \\
    &= s\cdot n_{s-1}-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+1)(s-2)}{2}
  \end{split}
\end{eqnarray}

Conventionally a given frequency constraint is met by selecting word-lengths of each stage of the CSA in order to satisfy~(\ref{TimingConstraint_CSA}).
\begin{eqnarray}\label{TimingConstraint_CSA}
  \begin{matrix}
    d_i\leqslant\frac{1}{f_s} &, & \forall i\in[0,s-1]
  \end{matrix}
\end{eqnarray}
Under this situation, the maximum word-length of the most significant stage $n_{s-1}$ is given by~(\ref{WL_MSB_CSA}), where $b$ is determined by~(\ref{Max.length of carry chain}).
\begin{eqnarray}\label{WL_MSB_CSA}
  n_{s-1}\cdot\mu_c+\mu_{mux}=b\cdot\mu_c
\end{eqnarray}

%, the word-lengths of RCAs and CSAs are selected to avoid timing violation, as presented in~(\ref{CSA_RCA}).

Substitute~(\ref{WL_MSB_CSA}) into (\ref{CSA_WL}), we derive the representation of the word-length of CSA in terms of a given timing constraint, as presented in~(\ref{CSA_Timing}).
% we form the relationship between the word-length of CSA and RCA under a given timing constraint, as presented in~(\ref{CSA_Timing}).
\begin{eqnarray}\label{CSA_Timing}
  n_{CSA}=s\cdot b-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+2)(s-1)}{2}
\end{eqnarray}

%It can be seen that $s=1$ leads to $n_{CSA}=n_{RCA}$. This is because RCA forms the least significant stage of CSA. In addition, combine (\ref{CSA_DelayRep}) and (\ref{CSA_WL}) to ensure $n_{s-1}>0$, we obtain the upper bond of the stage number by~(\ref{CSA_StageNumber}).
%\begin{eqnarray}\label{CSA_StageNumber}
%  s<n_{RCA}\cdot\frac{\mu_c}{\mu_{mux}}+1
%\end{eqnarray}

\subsection{Model Verification}
%As seen in (\ref{CSA_Timing}), the value of $\mu_{mux}/\mu_c$ should be determined before applying the timing model. This ratio can be obtained through timing analysis instead of experiments. Specifically, varying the word-length of a single stage $n_i$ and recording the corresponding delay value $d_i$ from the timing analysis tool. According to~(\ref{CSA_SingleStageDelay}), the ratio can be obtained by fitting those values using linear polynomials.

%Using this information, we verify our timing model with experimental results, which are obtained from post place and route simulations on Xilinx Virtex-6 FPGAs. Fig.~\ref{CSA Model Verification} illustrates both the modeled value and the experimental results of the maximum word-length of the 2-stage CSA for various operating frequencies. In the following experiments within this section, the input data are randomly sampled from a 16-bit data set, which follows uniform distribution. It can be seen that the the modeled value match well with the experimental results.
The ratio $\mu_{mux}/\mu_c$ can be computed from experiments. We perform post place-and-route simulations on the CSA with 2 stages using Xilinx Virtex-6 FPGA. The delay of the $i^{th}$ stage $d_i$ in~(\ref{CSA_SingleStageDelay}) is recorded with respect to different word-lengths of this given stage ($n_i$). Then the total word-length of CSA can be predicted through~(\ref{CSA_Timing}). In addition, the maximum word-length of the 2-stage CSA is obtained experimentally by increasing $n_{CSA}$ until errors are observed at the output. The comparison between the modeled value and the empirical results of $n_{CSA}$ is illustrated in Fig.~(\ref{CSA Model Verification}). It can be seen that our model match well with the experimental results.
%
%experimentally and compare our model with the empirical results of the maximum word-length of the 2stage CSA for various operating frequencies. It can be seen that the modeled value match well with the experimental results.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=3in]{./Figures/Model.eps}
  \caption{Comparison of the maximum word-length between the modeled value and the experimental results from FPGA simulations.}
  \label{CSA Model Verification}
\end{figure}

@@

\subsection{Accuracy Benefits and Area Overhead in CSA}
As greater word-length corresponds to smaller truncation errors, we compare the maximum word-length of RCA and CSA with 2 stages and 3 stages over a range of operating frequencies, as presented in Fig.~\ref{CSA3stage Timing}. It can be seen from Fig.~\ref{CSA3stage Timing} that in comparison to RCA, CSA achieves greater word-length when frequency is initially increased. RCA only outperforms than CSA when very high frequency is applied. This is because at low frequencies, although the multiplexer delay limits the word-length of each stage in CSA when compared to RCA, the stage parallelism in CSA enables a greater word-length. However when frequency increases, the multiplexer delay becomes comparable to the delay of the carry chain, and this inhibits the benefits of parallelism. In addition, we can see that the word-length of 3-stage CSA is always greater than 2-stage CSA across the entire frequency domain.
\begin{figure}[tbp]
    \begin{minipage}[b]{.48\textwidth}
        \centering
        \subfigure[Maximum word-length.]{
            \includegraphics[width=3.3in]{./Figures/CSA3stage_Timing.eps}
            \label{CSA3stage Timing}
        }
        \subfigure[Hardware resource usage.]{
            \includegraphics[width=3.3in]{./Figures/CSA3stage_Area.eps}
            \label{CSA3stage Area}
        }
    \end{minipage}
    \centering
    \caption{Comparison between RCA and CSA across a variety of frequency values in terms of the maximum word-length of input signal and the area consumption.}
\end{figure}

However, the accuracy benefits brought by CSA comes at the cost of a large area overhead. Fig.~\ref{CSA3stage Area} depicts the number of Look-Up Tables (LUTs) in the FPGA used for all three structures. It can be seen that in order to meet a given frequency, the 3-stage CSA consumes $2.4\times\sim3.7\times$ area than RCA, while the 2-stage CSA requires $1.7\times\sim3.1\times$ extra area. This finding poses a question of which is the best adder structure for a specific area budget.
%whether CSA with higher stage numbers still offer better accuracy than RCA under a limited area budget.
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/CSA3stage_Area.eps}
%  \caption{Area overhead for RCA and CSA.}
%  \label{CSA3stage Area}
%\end{figure}

\subsection{Exploring Trade-offs Between Accuracy, Performance and Area}
In the conventional design scenario, the word-length of the RCA and CSA are limited by the following two factors:
\begin{itemize}
  \item The critical path of the adder;
  \item The available silicon area.
\end{itemize}
We compare the accuracy, performance and area trade-offs to an RCA where the maximum possible word-length is limited by the given area budget, and the timing constraints are allowed to be violated.
%To answer this question, the silicon area is incorporated into the trade-offs together with accuracy and performance. If the available hardware resources are limited, the full word-length of both CSA and RCA might not be implemented. The precision loss might results in large errors even at low frequencies. This is demonstrated by experiments where area constraints are applied besides the timing requirement. In the conventional design scenario, both of the constraints are met by reducing the word-length of the input signal. The errors at the outputs are recorded with the reference to the original word-length, which is 16 bits in the following experiments. In addition, we propose another circuit designing method where RCA is implemented with the maximum possible word-length under the given area budget, while the timing constraints are met by overclocking.

%Our results could be of interest to a circuit designer in two aspects. On one hand, it is desirable to create a circuit that can run at a given frequency with the minimum achievable output errors and resource usage. On the other hand, the algorithm designer will wish the circuit to operate as fast as possible with the minimum area whilst a certain error budget can be tolerated. We select the optimum design methodology for both situations.

Suppose an algorithm designer would wish to create a circuit that can run at a given frequency with the minimum achievable output errors and resource usage. In this case, for a certain pair of $\{Area, Frequency\}$ constraint, the optimum design method is selected based on the following criteria:
%For instance, Fig.xxx to Fig.xxx illustrate error expectations of all the described design scenarios when setting the LUTs number to 45, 35, 25 and 15 respectively. The optimal design metric with the minimum error expectation is labelled for each area constraint. For a given frequency and area constraint, the optimum design is chosen based on the following metrics:
\begin{itemize}
  \item Design with the best accuracy at outputs is the optimum design;
  \item If multiple designs achieve the same accuracy, then the design with minimum area is the optimum design;
  \item If the accuracy and area are identical for multiple designs, they are all treated as the optimum design.
\end{itemize}

If we define the accuracy in terms of the error expectation at the output. The optimum design method with respect to a variety of operating frequencies and area consumptions is demonstrated in Fig.~\ref{Tradeoff}. From this figure several observations can be made. If the available area is large enough to implement a CSA in full precision, it will be the optimum design. This is expected from our earlier analysis in Fig~(\ref{CSA3stage Timing}). The 2-stage CSA is better than the 3-stage CSA when frequency is initially increased, as it consumes less area although both of them achieve the same error expectation. For a tighter area budget, only part of the CSA can be implemented, whereas the RCA still keeps full precision. In this case, area becomes the dominate factor and precision is lost for the CSA. In this case we see the RCA with overclocking is the optimum design method across almost the whole frequency domain. For more stringent area constraints, the word-length of RCA is also limited. This results in truncation error initially for all design scenarios. However, the RCA with overclocking is still .......
%Meanwhile, CSA with high stage numbers might not even be implemented at all.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=3.5in]{./Figures/Tradeoff.eps}
  \caption{Demonstration of the optimum design methodology which achieves the minimum error at outputs with respect to a variety of frequency and area constraints.}
  \label{Tradeoff}
\end{figure}

Similarly if the algorithm designer would expect the circuit to operate as fast as possible with the minimum area whilst a certain error budget can be tolerated, the optimum design methodology can be decided as illustrated in Fig.~(\ref{Tradeoff_Error}). In this situation, the error specifications are evaluated in terms of mean relative error (MRE), as presented in~(\ref{MRE}), where $E_{error}$ and $E_{out}$ refer to the mean value of error and the mean value of outputs, respectively.
\begin{eqnarray}\label{MRE}
  MRE=\left|\frac{E_{error}}{E_{out}}\right|\times 100\%
\end{eqnarray}

In our experiments, MRE is set ranging from $0.001\%$ to $50\%$. For a certain MRE, the design with the maximum operating frequencies is labeled as the optimum design. Moreover, the smallest design is the optimum one if multiple structures operate at the same frequency, with a certain area. Based on these criteria, the decision graph is depicted in Fig.~(\ref{Tradeoff_Error}).
%We also notice that for small frequencies, RCA with either overclocking or truncation of word-length of the input signal is the best design choice for any area constraint, since
\begin{figure}[htbp]
    \centering
    \includegraphics[width=3.5in]{./Figures/Tradeoff_Error.eps}
    \caption{Demonstration of the optimum design methodology which runs at the fastest frequency with respect to a variety of accuracy and area constraints.}
    \label{Tradeoff_Error}
\end{figure}

For a tight accuracy requirement, i.e. $MRE<0.005\%$, CSA serves as the optimum design choice with respect to large accessible area, as it intrinsically operates faster than RCA. Once again, when the area budget shrinks, the RCA performs best because the precision of the CSA is limited. Similarly to the previous section, we see that the overclocked RCA achieves the fastest operating frequencies with respect to most area constraints.

%the precision of RCA is reduced when critical area requirement is applied.
%
%However, if the tolerable error increases, i.e. $MRE>1\%$, the overclocked RCA achieves the fastest operating frequencies with respect to most area constraints.
